---
title: 'Classification Algorithm for Supervised Learning - Heart Disease Diagnosis'
author: 'Tyler Blakeley, Benjamin Kan, Mohammad Islam, Avijeet Singh'
date: 'October 12 2018'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Statement

Moe to complete

# Business Statements

Moe to complete

# Data Understanding

## Data Source and Collection

The data are collected from the UCI Machine Learning Repository site (https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/heart-disease.names). The datasets are originated from the four heart disease diagnosis databases from the following four locations dated July 1988:

* Hungarian Institute of Cardiology. Budapest: Andras Janosi, M.D.        
* University Hospital, Zurich, Switzerland: William Steinbrunn, M.D.        
* University Hospital, Basel, Switzerland: Matthias Pfisterer, M.D.        
* V.A. Medical Center, Long Beach and Cleveland Clinic Foundation: Robert Detrano, M.D., Ph.D.

The four datasets originally consist of 76 attributes. However, all of the experiments referred to a subset of 14 attributes in which the UCI site also provided with another version of the datasets with these 14 attributes. We took the liberty to use the 14-attribute datasets instead for our prediction exercise.

## Data Descriptions
We aggregated the four datasets described above to create a single dataset. The attribute names along with their descriptions are listed in the table below:

Variable Name | Description
--------------|-------------------------------------------------------------------------------------------
AGE           | Age in years
SEX           | Sex: (1 = male; 0 = female)
CP            | Chest Pain Type: (1:typical angina; 2:atypical angina; 3:non-anginal pain; 4:asymptomatic)
TRESTBPS      | Resting Blood Pressure (in mm Hg on admission to the hospital)
CHOL          | Serum Cholestoral in mg/dl
FBS           | Fasting Blood Sugar > 120 mg/dl)  (1 = true; 0 = false)
RESTECG       | Resting Electrocardiographic Results (0:normal; 1:having ST-T wave abnormality; 2:showing                      | probable or definite left ventricular hypertrophy) 
THALACH       | Maximum Heart Rate Achieved
EXANG         | Exercise Induced Angina (1 = yes; 0 = no)
OLDPEAK       | ST Depression Induced by Exercise Relative to Rest 
SLOPE         | The Slope of the Peak Exercise ST Segment (1:upsloping; 2:flat; 3:downsloping) 
CA            | Number of Major Vessels (0-3) Colored by Flourosopy
THAL          | 3 = normal; 6 = fixed defect; 7 = reversable defect
TARGET        | Diagnosis of Heart Disease (0: < 50% diameter narrowing; 1: > 50% diameter narrowing)

## Data Exploration

### Load Packages
```{r, message=FALSE}
#import packages;
library(dplyr)
library(reshape2)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(mice)
library(VIM)
library(pROC)
library(caret)
library(sqldf)
library(corrgram)
library(GGally)
library(ggthemes) 
```
### Load Datasets
We can now load in the four datasets and then we merge them into one.

```{r, message=FALSE}

data_cleveland <- read.csv(file.choose(),header = TRUE, na.strings = c("NA","","#NA","?"))
data_hungarian <- read.csv(file.choose(), header = TRUE, na.strings = c("NA","","#NA","?"))
data_switzerland <- read.csv(file.choose(), header = TRUE, na.strings = c("NA","","#NA","?"))
data_VA <- read.csv(file.choose(), header = TRUE, na.strings = c("NA","","#NA","?"))
full_data <- rbind(data_cleveland,data_hungarian,data_switzerland,data_VA)
```
Upon initial inspection of the data, we found that the data convention for the TARGET attribute is inconsistent among the 4 datasets. Both Switzerland and VA datasets have the heart disease dignosis target variable with the values of 0, 1, 2, 3 and 4 instead of the values of 0 or 1. Here we will make the data format consistent with the data definition described above (i.e 0: < 50% diameter narrowing; 1: > 50% diameter narrowing). So we made the following adjustments on the TARGET attribute:
```{r, message=FALSE}
# original target variable distributions
table(full_data$TARGET)
full_data$TARGET <- ifelse(full_data$TARGET>=1,1,full_data$TARGET)
# final target variable distributions
table(full_data$TARGET)
```


